---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Accessing Data

```{r cars}

cm <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv")
head(cm)
data("cm")

library(dplyr)
library(ggplot2)
```

```{r}
#Cleaning
cm <- cm %>%
  filter(!is.na(Total), Total > 0)

cm <- cm %>%
  mutate(
    ShareMen = Men / Total,
    EmploymentRate = Employed / Total,
    UnemploymentRate = Unemployed / (Employed + Unemployed)
  )

head(cm)
```
##Introduction

Often time there are assumptions surrounding college majors. What majors pay better, what will get you a job, and what majors women tend to gravitate towards. Being college students, we see stereotypes and assumptions like this everywhere, but are any of them actually true?

##Data

This is data from the fivethirtyeight package on github. It explores various college majors, their employment rates, average income, and percentage of graduates that are women. The data was taken from the American Community Survey in 2010-2012. 

## Questions

What major category has the highest percentage of full time employment after graduation?

Do major categories with higher percentages of women have higher employment rates?

Which majors provide the best balance of employment rate and median salary?

## Results

What major category has the highest percentage of full time employment after graduation?

```{r}
cm <- cm %>%
  mutate(FullTimeShare = Full_time / Employed)

cm <- cm %>%
  filter(!is.na(FullTimeShare), Employed > 0)

category_summary <- cm %>%
  group_by(Major_category) %>%
  summarize(AvgFullTimeShare = mean(FullTimeShare, na.rm = TRUE)) %>%
  arrange(desc(AvgFullTimeShare))

#Average Full time employment by major type
ggplot(category_summary, aes(x = reorder(Major_category, AvgFullTimeShare), y = AvgFullTimeShare, fill = Major_category)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Average Full-Time Employment Share by Major Category",
    x = "Major Category",
    y = "Average Full-Time Employment Share"
  ) +
  scale_y_continuous(labels = scales::percent_format())

#Faceted by major
ggplot(cm, aes(x = reorder(Major, FullTimeShare), y = FullTimeShare)) +
  geom_col(fill = "steelblue") +
  facet_wrap(~ Major_category, scales = "free_y") +
  labs(
    title = "Full-Time Employment Share by Major (Faceted by Major Category)",
    x = "Major",
    y = "Full-Time Employment Share"
  ) +
  scale_y_continuous(labels = scales::percent_format())

#Major colored to major type
ggplot(cm, aes(x = reorder(Major, FullTimeShare), y = FullTimeShare, color = Major_category)) +
  geom_col() +
  labs(
    title = "Full-Time Employment Share by Major",
    x = "Major",
    y = "Full-Time Employment Share"
  ) +
  scale_y_continuous(labels = scales::percent_format())

```




Do major categories with higher percentages of women have higher employment rates?

```{r}
#Filter
cm <- cm %>%
  filter(Total > 0) %>%
  mutate(
    ShareWomen = Women / Total,
    EmploymentRate = Employed / Total
  )

# Aggregate by major category
category_summary <- cm %>%
  group_by(Major_category) %>%
  summarize(
    AvgShareWomen = mean(ShareWomen, na.rm = TRUE),
    AvgEmploymentRate = mean(EmploymentRate, na.rm = TRUE)
  )

# Plot: Share of Women vs Employment Rate
ggplot(category_summary, aes(x = AvgShareWomen, y = AvgEmploymentRate, label = Major_category)) +
  geom_point(, size = 4) +
  geom_text(nudge_y = 0.01, size = 3) +
  labs(
    title = "Women and Employment Rates", 
    x = "Average % Women (per Category)",
    y = "Average Employment Rate"
  ) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_y_continuous(labels = scales::percent_format())

ggplot(cm, aes(x = ShareWomen, y = EmploymentRate)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Employment Rate vs Share of Women", x = "Share of Women", y = "Employment Rate") +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal()

#Women by Salary
library(scales)

ggplot(cm, aes(x = ShareWomen, y = Median, color = Major_category)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Median Salary vs Share of Women by Major", x = "Share of Women", y = "Median Salary") +
  scale_x_continuous(labels = scales::percent_format()) + 
  scale_y_continuous(labels = label_dollar(scale = 0.001, suffix = "K")) + 
  theme(legend.position = "bottom")

#Men by Salary

ggplot(cm, aes(x = ShareMen, y = Median, color = Major_category)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Median Salary vs Share of Men by Major", x = "Share of Men", y = "Median Salary") +
  scale_x_continuous(labels = scales::percent_format()) + 
  scale_y_continuous(labels = label_dollar(scale = 0.001, suffix = "K")) + 
  theme(legend.position = "bottom")

#Pie Chart
women_by_category <- cm %>%
  group_by(Major_category) %>%
  summarize(TotalWomen = sum(Women, na.rm = TRUE)) %>%
  mutate(Percent = TotalWomen / sum(TotalWomen)) %>%
  arrange(desc(Percent)) %>%
  mutate(label_pos = cumsum(Percent) - Percent / 2)

women_by_category


ggplot(women_by_category, aes(x = "", y = TotalWomen, fill = Major_category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(
    title = "Distribution of Women by Major Category",
    x = NULL,
    y = NULL,
    fill = "Major Category"
  ) +
  theme_void() +
  theme(legend.position = "right")


#Male Pie Chart
men_by_category <- cm %>%
  group_by(Major_category) %>%
  summarize(TotalMen = sum(Men, na.rm = TRUE)) %>%
  mutate(Percent = TotalMen / sum(TotalMen)) %>%
  arrange(desc(Percent)) %>%
  mutate(label_pos = cumsum(Percent) - Percent / 2)

men_by_category


ggplot(men_by_category, aes(x = "", y = TotalMen, fill = Major_category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(
    title = "Distribution of Men by Major Category",
    x = NULL,
    y = NULL,
    fill = "Major Category"
  ) +
  theme_void() +
  theme(legend.position = "right")
```

Which majors provide the best balance of employment rate and median salary?
```{r}
library(dplyr)
library(ggplot2)

# Load and prepare the data
cm <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/recent-grads.csv")

# Create EmploymentRate column
cm <- cm %>%
  mutate(EmploymentRate = Employed / Total) %>%
  filter(!is.na(EmploymentRate), !is.na(Median))

# Rank by combined employment rate and median salary (simple sum for ranking)
cm$Score <- scale(cm$EmploymentRate) + scale(cm$Median)

top_10 <- cm %>%
  arrange(desc(Score)) %>%
  slice(1:10)

# Bar plot
ggplot(top_10, aes(x = reorder(Major, Median), y = Median, fill = Major_category)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 10 Majors with Best Balance of Employment & Salary",
    subtitle = "Ranked by combined employment rate and median salary",
    x = "Major",
    y = "Median Salary (USD)",
    fill = "Major Category"
  )

ggplot(cm, aes(x = EmploymentRate, y = Median, color = Major_category)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_hline(yintercept = mean(cm$Median, na.rm = TRUE), linetype = "dashed") +
  labs(
    title = "Employment Rate vs. Median Salary by Major",
    x = "Employment Rate",
    y = "Median Salary (USD)",
    color = "Major Category"
  )

```

This bar chart shows the ten college majors that offer the best mix of earning potential and job security. It ranks majors based on a combined score of median salary and employment rate, helping highlight which fields give you the best overall value. Unsurprisingly, engineering-related majors like Petroleum and Architectural Engineering top the list. The color coding breaks the majors into categories like Business, Engineering, or Arts, making it easy to see which areas tend to perform better. It’s a helpful chart for students thinking about what major can lead to both a good paycheck and a solid chance of getting hired.



This scatter plot maps out how all college majors compare in terms of employment rate and salary. Each dot is a major, with salary on the vertical axis and employment rate on the horizontal. The dotted line shows the average salary overall, which makes it easy to spot which majors are above or below that line. The colors group majors into categories like Health, Business, Engineering, and so on. It’s a quick way to see which fields give you the best shot at a job, which pay the most, and which do both. Engineering majors, for example, usually land in the top-right — high pay and high employment.




##Conclusions
Conclusions



1. What major category has the highest percentage of full-time employment after graduation?

Our analysis found that engineering and health-related majors tend to have the highest share of full-time employment after graduation. These fields consistently lead to stable, full-time positions, likely due to strong demand and the technical skills associated with these majors. This suggests that students pursuing these degrees are more likely to secure reliable employment shortly after completing their studies.

2. Do major categories with higher percentages of women have higher employment rates?

We observed that majors with a higher proportion of women, such as education and social work, often have lower average employment rates compared to more male-dominated fields like engineering or computer science. This doesn’t imply causation, but it may reflect broader labor market dynamics, such as wage disparities or differences in industry demand. These findings highlight ongoing gender-related challenges in the workforce that are worth exploring further.

3. Which majors provide the best balance of employment rate and median salary?

Choosing a college major is one of the most important decisions students make, and our analysis shows just how much that choice can affect future job prospects and earning potential. By analyzing employment rates and median salaries across a wide range of majors, we found that some fields particularly in engineering and computer science consistently offer both strong job security and higher pay. On the other hand, many majors in the arts and humanities tend to offer lower salaries and more limited job prospects, though they may still provide value in less measurable ways. Ultimately, while passion and interest should always play a role in selecting a major, having clear data on economic outcomes can help students make more informed and confident decisions about their education and career paths.




